---
title: "K-MEANS"
author: "rajendra"
date: "10/31/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r}
library(readr)
library(caret)
library(ISLR)
library(tidyverse)
library(factoextra)
set.seed(15)
```


Qa :- Eliminating null values from the dataset
```{r}
# Importing dataset in r
Universities <- read_csv("C:/Users/rajendra/Downloads/Universities.csv")
uni<-na.omit(Universities)
summary(uni)

```


Qb :- Obtaining the opitmal value of k and number of clusters

```{r}
# Removing Categorical values from the dataset

a<-uni[,c(-1,-2,-3)]

#Scaling  the data frame (z-score)

b<-scale(a)

summary(b)


#Finding the best k value using Elbow Chart

fviz_nbclust(b,kmeans,method = "wss")

#Using silhouette method to determine k

fviz_nbclust(b,kmeans,method = "silhouette")

#Kmeans cluster algorithm
k4 <- kmeans(b, centers = 3, nstart = 1)

k4$centers # Centroids

k4$size    # Size of each cluster

# Visualizing the clusters

fviz_cluster(k4,data=b)


```

From the graph,we can say that k=3 is the best value and the dataset is divided into three different  clusters

Qc :- Combining cluster information with dataset to determine the statistics of each cluster
```{r}
com<-cbind(uni,k4$cluster) 

# Assigning labels to three clusters (Tier 1, Tier 2, Tier3 Universities)

com$`k4$cluster`<-factor(com$`k4$cluster`,levels = c(1,2,3),labels = c("T3","T1","T2")) 

c1<-com[com$`k4$cluster`=="T1",]#Cluster 2

c2<-com[com$`k4$cluster`=="T2",]#Cluster 3

c3<-com[com$`k4$cluster`=="T3",]#Cluster 1



plot(c(0), xaxt = 'n', ylab = "", type = "l",
     ylim = c(min(k4$centers), max(k4$centers)), xlim = c(0, 18))

axis(1, at = c(1:17), labels = FALSE)
text(seq(1,17,by=1),par("usr")[3]-0.2,labels = colnames(b),srt = 45,pos =1,xpd = TRUE) # X Measurement names

# plot centroids
for (i in c(1:3))
  lines(k4$centers[i,], lty = i, lwd = 2) 

# name clusters
text(x = 0.5, y = k4$centers[, 1], labels = paste("Cluster", c(1:3))) # Cluster Names


```

Classifying the clusters based on their mean values of respective columns

Cluster 1 Statistics
```{r}
#From the graph shown above we can infer the characterstics of cluster 1 and named it as Tier 3 Universities

#Tier 3 universties (Cluster 1)

#Low applications recieved
#Low applications accepted
#Minimum number of students enrolled
# Minimum number of new students from top 25%
# Minimum number of new students from top 10%
# Less number of full time under graduate
# Out of station tution fee are less expensive
# Very few number of rooms
#Low cost of books
# Number of boards are low
# Graduation rate is low
# less percent of faculty ratio wrt phd



```
Cluster 2 statistics
```{r}
#From the graph shown above we can infer the characterstics of cluster 2 and named it as Tier 1 Universities

# Tier 1 Universties (Cluster 2)
# Higher Graduation rate             
# High percentage faculty ratio w.r.t PHD   
# High in-state tutuion fee
# More number of rooms             
# Number of boards are high        
# More number of new students from top 25%
# More number of new students from top 10%

```
cluster 3 statistics
```{r}
#From the graph shown above we can infer the characterstics of cluster 3 and named it as Tier 2 Universities

#Tier 2 universties (Cluster 3)
#Maximum number of applications recieved    
#Maximum number of applications accepted    
#Maximum number of new students enrolled
#More number of part time graduates
#More number of full time graduates
#Huge amount of estimated book cost
#Huge amount of personal expenses
#Higher Student to faculty ratio

```

Qd :- Finding the relationship between the clusters and the categorical information?

```{r}
head(com)
head(com[com$`k4$cluster`=="T1",c(1,2,3,21)]) 
head(com[com$`k4$cluster`=="T2",c(1,2,3,21)]) 
head(com[com$`k4$cluster`=="T3",c(1,2,3,21)])
library(ggplot2)
ggplot(com,aes(x=com$State,y=com$`Public (1)/ Private (2)`,color=com$`k4$cluster`))+geom_point()

```


cluster 1:-It has both public and private universities
cluster 2:- maximum number of the universities are private
cluster 3::-maximum number of the universities are pubic

Qe :-
```{r}
k4
# 1st cluster has 150 colleges and has a variability of 1424.8 

# 2nd cluster has 46 colleges  and has a variability of 1044.68 (the members are closer to each other in terms of distance hen compared to other clusters)

# 3rd cluster has 275 colleges  and has a variability of 2562.3
k4$withinss
# It describes the distance between centroids and mean of all the points within a cluster

k4$betweenss
# Itdescribes the distance between centriods of all the customers
```

Qf :-
```{r}


Km<-kmeans(a,centers = 3)
b1<-mean(Km$centers[1,]) # Mean of Cluster 1
b2<-mean(Km$centers[2,]) # Mean of cluster 2
b3<-mean(Km$centers[3,]) # Mean of cluster 3
a1<-Universities[Universities$`College Name`=="Tufts University",]
View(a1)
a2<-apply(a1[,-c(1:3,10)],1,mean) # Mean of record
dist(rbind(a2,b1)) # Euclideam distance betweewn  cluster 1 mean and Tufts university data
dist(rbind(a2,b2))
dist(rbind(a2,b3))
a1$`# PT undergrad`<-1596.33 # From the above, Mean value which is near to cluster 2. Hence replacing the missing value with mean value
uni2<-rbind(uni,a1)
View(uni2)
uni2_z<-scale(uni2[,-c(1:3)])
uni2_cluster<-kmeans(uni2_z,3)
uni2<-cbind(uni2,uni2_cluster$cluster)
uni2[472,] # From the model, this uniersity falls under Cluster 2

```

Tufts University belongs to cluster 2 from the above information and it belongs to Tier 1 University
