---
title: "K-MEANS"
author: "rajendra"
date: "10/31/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r}
library(readr)
library(caret)
library(ISLR)
library(tidyverse)
library(factoextra)
set.seed(15)
```


Qa :- Eliminating null values from the dataset
```{r}
# Importing dataset in r
Universities <- read_csv("C:/Users/rajendra/Downloads/Universities.csv")
uni<-na.omit(Universities)
summary(uni)

```


Qb :- Obtaining the opitmal value of k and number of clusters

```{r}
# Removing Categorical values from the dataset

a<-uni[,c(-1,-2,-3)]

#Scaling  the data frame (z-score)

b<-scale(a)

summary(b)


#Finding the best k value using Elbow Chart

fviz_nbclust(b,kmeans,method = "wss")

#Using silhouette method to determine k

fviz_nbclust(b,kmeans,method = "silhouette")

#Kmeans cluster algorithm
k4 <- kmeans(b, centers = 3, nstart = 1)

k4$centers # Centroids

k4$size    # Size of each cluster

# Visualizing the clusters

fviz_cluster(k4,data=b)


```

From the graph,we can say that k=3 is the best value and the dataset is divided into three different  clusters

Qc :- Combining cluster information with dataset to determine the statistics of each cluster
```{r}
com<-cbind(uni,k4$cluster) 

# Assigning labels to three clusters (Tier 1, Tier 2, Tier3 Universities)

com$`k4$cluster`<-factor(com$`k4$cluster`,levels = c(1,2,3),labels = c("T3","T1","T2")) 

c1<-com[com$`k4$cluster`=="T1",]#Cluster 2

c2<-com[com$`k4$cluster`=="T2",]#Cluster 3

c3<-com[com$`k4$cluster`=="T3",]#Cluster 1


```

Classifying the clusters based on their mean values of respective columns

Cluster 1 Statistics
```{r}
summary(c1)

#Tier 3 universties (Cluster 1)

#Low applications recieved
#Low applications accepted
#Minimum number of students enrolled
# Minimum number of new students from top 25%
# Minimum number of new students from top 10%
# Less number of full time under graduate
# Out of station tution fee are less expensive
# Very few number of rooms
#Low cost of books
# Number of boards are low
# Graduation rate is low
# less percent of faculty ratio wrt phd



```
Cluster 2 statistics
```{r}
summary(c2)
# Tier 1 Universties (Cluster 2)
# Higher Graduation rate             
# High percentage faculty ratio w.r.t PHD   
# High in-state tutuion fee
# More number of rooms             
# Number of boards are high        
# More number of new students from top 25%
# More number of new students from top 10%

```
cluster 3 statistics
```{r}
summary(c3)
#Tier 2 universties (Cluster 3)
#Maximum number of applications recieved    
#Maximum number of applications accepted    
#Maximum number of new students enrolled
#More number of part time graduates
#More number of full time graduates
#Huge amount of estimated book cost
#Huge amount of personal expenses
#Higher Student to faculty ratio

```

Qd :- Finding the relationship between the clusters and the categorical information?

```{r}
head(com)
head(com[com$`k4$cluster`=="T1",c(1,2,3,21)]) 
head(com[com$`k4$cluster`=="T2",c(1,2,3,21)]) 
head(com[com$`k4$cluster`=="T3",c(1,2,3,21)])  
```


cluster 1:-It has both public and private universities
cluster 2:- maximum number of the universities are private
cluster 3::-maximum number of the universities are pubic

Qe :-
```{r}
k4
# 1st cluster has 150 colleges and has a variability of 1424.8 

# 2nd cluster has 46 colleges  and has a variability of 1044.68 (the members are closer to each other in terms of distance hen compared to other clusters)

# 3rd cluster has 275 colleges  and has a variability of 2562.3
k4$withinss
# It describes the distance between centroids and mean of all the points within a cluster

k4$betweenss
```

Qf :-
```{r}


Universities$`# PT undergrad`[is.na(Universities$`# PT undergrad`)]=mean(Universities$`# PT undergrad`,na.rm=TRUE)            # Replacining the na values with mean of that column
test<-filter(Universities,`College Name` == "Tufts University") # filtering the Tufts university 
summary(Universities$`# PT undergrad`)

a1<-rbind(uni,test)


b1<-scale(a1[,-c(1,2,3)]) #Scaling  the data frame (z-score)


k5 <- kmeans(b1, centers = 3, nstart = 10) # kmeans algorithim

which(grepl("Tufts University",a1$`College Name`)) # to find the position of tufts university
k5$cluster[472] # To find the cluster


```

Tufts University belongs to cluster 2 from the above information and it belongs to Tier 1 University