---
title: "bathsoap"
author: "rajendra"
date: "12/5/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r}
library(caret)
library(factoextra)
library(hrbrthemes)
library(GGally)
library(viridis)
set.seed(123)

#Read Data
library(readr)

BathSoap <- read_csv("C:/Users/rajendra/Downloads/BathSoap.csv")
str(BathSoap)

#Customer Brand loyality
#Brand loyality is defined as the  customer buys spends a maximum amount of money in 8 brands.

r1<-BathSoap[,23:30]# Inclding 8 brands
BathSoap$Loyality_Brand<-as.numeric(apply(r1,1,max)) # Maximum value of the brand
table(BathSoap$Loyality_Brand)
```

a. The variables that describe purchase behavior (including brand loyalty)
```{r}
#The variables used by us are No.Of Brands,Brand runs, Total Volume,Number of Transctions,value,trans/brandruns,vol/trans,avg price,others999,Loyality_Brand.

BS<-BathSoap[,c(12,13,14,15,16,17,18,19,31,47)]

data1.s<-as.data.frame(scale(BS)) # scaling the data
# Elbow chart to estimate the optimal K
fviz_nbclust(data1.s,kmeans,method = "silhouette")

# Choosing the optimal K as 3 and forming 3 clusters
model<-kmeans(data1.s,3,nstart=50)

# Visualizing the clusters
fviz_cluster(model,data1.s)


result<-as.data.frame(cbind(1:nrow(model$centers),model$centers))
result$V1<-as.factor(result$V1)
# Characteristics of the cluster
result

model$size


# Parallel plot to visualize the cluster.
ggparcoord(result,
           columns = 2:ncol(result), groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
) 


```
Characterstics of cluster based on purchase behaviour

Loyality brand for cluster1 is very high and cluster 2 is very low because the number of brand runs is very low and high respctively.

Average price is low for cluster 1 and high for cluster 2 and moreover the people are purchasing very low volume of brands from others 999 in cluster 1  when compared to cluster 2

High volume of transctions is very high in clster1 when compared to cluster 2.

Cluster 3 is unremarkable from any measure
 

b. The variables that describe the basis for purchase

```{r}
#Finding the maximum proptional category from the 10 categories
r2<-BathSoap[,36:46]
BathSoap$max_prop_no<-as.numeric(apply(r2,1,which.max))
BathSoap$max_prop<-as.numeric(apply(r2,1,max))

# variables considered for basis of purchase are purchase volume no promo,6%,other promo,pric categories from 1 to 4 and bath soap maximumprop
BS1<-BathSoap[,c(20:22,32:35,49)]

data2.s<-as.data.frame(scale(BS1)) # scaling the data
# Elbow chart to estimate the optimal K
fviz_nbclust(data2.s,kmeans,method = "silhouette")

# Choosing the optimal K as 3 and forming 3 clusters
model1<-kmeans(data2.s,3,nstart=50)

# Visualizing the clusters
fviz_cluster(model1,data2.s)


result1<-as.data.frame(cbind(1:nrow(model1$centers),model1$centers))
result1$V1<-as.factor(result1$V1)
# Characteristics of the cluster
result1

model1$size


# Parallel plot to visualize the cluster.
ggparcoord(result1,
           columns = 2:ncol(result1), groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
) 
```
Characterstics of cluster

In cluster 1 we can see that maximum proption of purchase is coming when  there is no promos and exactly  reversible for cluster 2

There is low for value for price category 1,category2,category 4 and high value of price category 3 then high chance of maximum proption of purchase in cluster1 


c:

```{r}

BS2<-BathSoap[,c(12:22,31:35,47,49)]

data3.s<-as.data.frame(scale(BS2)) # scaling the data
# Elbow chart to estimate the optimal K
fviz_nbclust(data3.s,kmeans,method = "silhouette")

# Choosing the optimal K as 3 and forming 3 clusters
model2<-kmeans(data3.s,3,nstart=50)

# Visualizing the clusters
fviz_cluster(model2,data3.s)


result2<-as.data.frame(cbind(1:nrow(model2$centers),model2$centers))
result2$V1<-as.factor(result2$V1)
# Characteristics of the cluster
result2

model2$size


# Parallel plot to visualize the cluster.
ggparcoord(result2,
           columns = 2:10, groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
) 
ggparcoord(result2,
           columns = 11:19, groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
) 
```



2)
```{r}
BS1<-BathSoap[,c(2:11,20:22,31:35,47,49)]

data2.s<-as.data.frame(scale(BS1)) # scaling the data
# Elbow chart to estimate the optimal K
fviz_nbclust(data2.s,kmeans,method = "silhouette")

# Choosing the optimal K as 3 and forming 3 clusters
model1<-kmeans(data2.s,3,nstart=50)

# Visualizing the clusters
fviz_cluster(model1,data2.s)


result1<-as.data.frame(cbind(1:nrow(model1$centers),model1$centers))
result1$V1<-as.factor(result1$V1)
# Characteristics of the cluster
result1

model1$size


# Parallel plot to visualize the cluster.
ggparcoord(result1,
           columns = 2:21, groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
) 

ggparcoord(result1,
           columns = c(2,10:14,19:21), groupColumn = 1, 
           showPoints = TRUE, 
           title = "Characterisitcs of the cluster",
           alphaLines = 0.3 
)
```


When we compare all the three models it is observed that variables with reasons of purchase  can explain all the characterstics of data when it is compared with purchase bhaviour and combination of both. so the best segementation of the model is reasons of purchase.

3)
```{r}
r1<-BathSoap[,23:31]
BathSoap$Loyality<-as.numeric(apply(r1,1,which.max))
BS3 <- BathSoap[,c(2:4,6:11,19,20:22,31:35,47,48,50)]
BS3$clusters <- model1$cluster
ggplot(BS3) +
  aes(x = clusters,fill=as.factor(HS)) +
  geom_bar() +
  scale_fill_hue() +
  theme_minimal() +
  facet_wrap(vars(c("Pr Cat 1","Pr Cat 2","Pr Cat 3","Pr Cat 4")))


ggplot(BS3) +
  aes(x = clusters,fill= as.factor(apply(BS3[,c(15:18)],1,which.max)))+
  geom_bar() +
  scale_fill_hue() +
  theme_minimal() +
  facet_wrap(vars(c("Pur Vol No Promo - %","Pur Vol Promo 6 %","Pur Vol Other Promo %")))


```
Suggested mail promotions
when it comes in cluster 1  there is a minimum purchase of pricecategory 4 and price category 1
even though where there is availabilty of all promos or no promos.

when it comes in cluster 2  there is a minimum purchase of pricecategory 3
even though where there is availabilty of all promos or no promos.

when it comes in cluster 3  there is a minimum purchase of pricecategory 4
even though where there is availabilty of all promos or no promos.



